---
title: "ZadokPMFWheat"
---


```{r setup, echo = FALSE}
setup_qmd()
```

The `ZadokPMFWheat` model calculates Zadoks growth stage values [@zadoks_decimal_1974] for [wheat](/docs/Plants/Wheat/Phenology/index.qmd) based on the current phenological growth stage within the simulation. This model provides a standardized way to express wheat development stages using the internationally recognized Zadok scale.

## Overview

The Zadoks scale [@zadoks_decimal_1974] is a numerical system (0-100) used to describe the growth stages of cereal crops, particularly wheat. The `ZadokPMFWheat` model translates APSIM's internal phenological stages into equivalent Zadok values, enabling researchers to communicate development stages using familiar agricultural terminology.



## Inputs

The following input functions are required by `ZadokPMFWheat`:

- `TillerNumber` — Total number of tillers the plant will develop.
- `HaunStage` — The current Haun stage of the plant.

## Methodology

The `ZadokPMFWheat` model calculates the Zadoks growth stage ($Z$) based on the current phenological phase of the wheat plant. Different calculation methods are applied depending on the developmental phase:

In **[Germinating](/docs/Plants/Wheat/Phenology/Germinating.qmd)**, Zadoks Stage ($Z$) is 5 times the fraction of the germination phase completed ($F(t)$):

$$
Z = 5.0 \times F(t)
$$

In **[Emerging](/docs/Plants/Wheat/Phenology/Emerging.qmd)**, Zadoks Stage ($Z$) is calculated as 5 plus 5 times the fraction of the emergence phase completed ($F(t)$):
$$
Z = 5.0 + (5.0 \times F(t))
$$

Before the phenological stage is less than 5.3 during [StemElongating](/docs/Plants/Wheat/Phenology/StemElongating.qmd) phase, Zadoks Stage ($Z$) is calculated as 10 plus the Haun stage value ($HS$):

$$
Z = 10.0 + HS
$$

After the phenological stage reaches 5.3, Zadoks Stage ($Z$) is determined through linear interpolation between predefined growth stages and their corresponding Zadok values:

| APSIM Growth Stage | Zadok Stage | Growth Stage Description | 
|-------------------|-------------|--------------------------|
| 5.0 | 30.0 | MaximumSpikeletPrimordia |
| 5.99 | 34.0 | Fourth node detectable |
| 6.0 | 39.0 | FlagLeafAppearance |
| 7.0 | 55.0 | HeadEmergence |
| 8.0 | 65.0 | Anthesis |
| 9.0 | 71.0 | MaximumGrainLength |
| 10.0 | 87.0 | EndGrainFill |
| 11.0 | 90.0 | HarvestRipe |

```{r fig-zadok-mapping, fig.height = 5, fig.cap="APSIM to Zadok Growth Stage Mapping"}
stage_data <- data.frame(
    APSIM_Stage = c(5.0, 5.99, 6.0, 7.0, 8.0, 9.0, 10.0, 11.0),
    Zadok_Stage = c(30.0, 34.0, 39.0, 55.0, 65.0, 71.0, 87.0, 90.0),
    Description = c("Terminal spikelet stage", "Beginning of stem elongation", 
                                    "Flag leaf stage", "Heading", "Anthesis", 
                                    "Kernel watery ripe", "Hard dough", "Harvest ripe")
)

# Create the plot
library(ggplot2)

ggplot(stage_data, aes(x = APSIM_Stage, y = Zadok_Stage)) +
    geom_line(color = "blue", size = 1.2) +
    geom_point(color = "red", size = 3) +
    geom_text(aes(label = paste0("Z", Zadok_Stage)), 
                        hjust = -0.1, vjust = -0.5, size = 3) +
    labs(title = "APSIM to Zadok Growth Stage Mapping",
             x = "APSIM Growth Stage",
             y = "Zadok Growth Stage",
             subtitle = "Linear interpolation between defined growth stages") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
                plot.subtitle = element_text(hjust = 0.5)) +
    scale_x_continuous(breaks = stage_data$APSIM_Stage) +
    scale_y_continuous(breaks = stage_data$Zadok_Stage)
```

At **HarvestRipe** stage, Zadok stage is fixed at 90.0:



## Events 

> No events are listened for by this function.

**Events Raised to**

> No events are raised by this function.

## Properties

> No configurable properties are available for this function.

**Read-Only Reportable Properties**

| Property             | Type    | Description                                                                 |
|----------------------|---------|-----------------------------------------------------------------------------|
| Stage     | double  | Zadok stage (Z) value                                                        |

## User Interface

> No user interface is provided for this function.

## Practical Example

> No practical example is available for this function.

## See Also

* **Source code:** [ZadokPMFWheat.cs on GitHub](http://github.com/APSIMInitiative/ApsimX/blob/master/Models/PMF/Phenology/Scales/ZadokPMFWheat.cs)
