---
title: "FrostHeatDamageFunctions"
format: html
---


```{r setup, echo = FALSE}
setup_qmd()
```

This `FrostHeatDamageFunctions` function computes daily and cumulative damage ratios (0–1) for frost and heat by combining a temperature‑driven potential yield reduction (piecewise linear between lower/upper thresholds with a maximum reduction) and a growth‑stage sensitivity factor (piecewise linear). The result represents the expected fractional yield loss from extreme temperature events.

## Overview


The `FrostHeatDamageFunctions` function is developed to account for the effects of frost and heat stresses on yield reduction, which runs at the daily step separating with yield estimation. The damage function is the product of the potential yield reduction induced by extreme(i.e.frost and heat) events and the sensitivity of the yield reduction to the growth stage when the events occur.The potential ratio of yield reduction induced by a frost or heat event is a piece-wise linear function of daily minimum or maximum air temperature, respectively.The potential ratio of yield reduction ranges from 0 to 1 indicating mild to severe yield reduction induced by an extreme event. The function is described with three parameters including the lower and upper temperature thresholds and the maximum yield reduction. The sensitivity of yield reduction to the growth stage is a piece-wise linear function of the growth stage simulated by APSIM. The sensitivity ranges from 0 to 1 indicating the least to most sensitivity of yield reduction.The function has four parameters: the lower and upper growth stage thresholds of the sensitive period to frost or heat stress, and the lower and upper growth stage thresholds of the most sensitive period around flowering when sensitivity equals 1. The same function of sensitivity applies to both frost and heat stress but with different parameterizations. Multiple frost and heat events during the sensitive period have multiplicative effects on yield reduction rather than additive effects.


## Inputs


The following inputs are required for the `FrostHeatDamageFunctions` function with default values for wheat and canola crops:

### Frost parameters

| Parameter (property) | Description | Wheat | Canola |
|---|---:|---:|---:|
| FrostLowTT | Lower temperature threshold for frost damage (°C) | -4.00 | -4.11 |
| FrostMaxReductionRatio | Yield reduction ratio at the lower threshold (0–1, max reduction) | 0.40 | 0.31 |
| FrostUpTT | Upper temperature threshold for frost damage (°C) | 1.00 | -0.75 |
| FrostMinReductionRatio | Yield reduction ratio at the upper threshold (0–1, min reduction) | 0.00 | 0.00 |
| FrostStartSensitiveGS | Start of the frost-sensitive growth stage period (GS units) | 6.49 | 7.22 |
| FrostStartMostSensitiveGS | Start of the most-sensitive period (sensitivity = 1) (GS units) | 8.00 | 7.33 |
| FrostEndMostSensitiveGS | End of the most-sensitive period (sensitivity = 1) (GS units) | 9.50 | 8.06 |
| FrostEndSensitiveGS | End of the frost-sensitive growth stage period (GS units) | 9.50 | 8.18 |

### Heat parameters

| Parameter (property) | Description | Wheat | Canola |
|---|---:|---:|---:|
| HeatLowTT | Lower temperature threshold for heat damage (°C) | 30.35 | 27.45 |
| HeatMinReductionRatio | Yield reduction ratio at the lower threshold (0–1, min reduction) | 0.00 | 0.00 |
| HeatUpTT | Upper temperature threshold for heat damage (°C) | 34.00 | 38.84 |
| HeatMaxReductionRatio | Yield reduction ratio at the upper threshold (0–1, max reduction) | 0.20 | 0.58 |
| HeatStartSensitiveGS | Start of the heat-sensitive growth stage period (GS units) | 7.28 | 5.58 |
| HeatStartMostSensitiveGS | Start of the most-sensitive period (sensitivity = 1) (GS units) | 7.48 | 6.26 |
| HeatEndMostSensitiveGS | End of the most-sensitive period (sensitivity = 1) (GS units) | 8.97 | 9.20 |
| HeatEndSensitiveGS | End of the heat-sensitive growth stage period (GS units) | 9.09 | 9.43 |

* **Reduction ratio** from 0 to 1 indicates no to full yield reduction,
* **Sensitivity** from 0 to 1 indicates no to full sensitivity to frost/heat damage.

The figures below illustrate the temperature response and growth stage sensitivity functions for frost and heat damage for both wheat and canola crops using the default parameters.

```{r example, fig.width=8, fig.height=6}
library(ggplot2)
library(patchwork)


# default parameters from selection
params <- list(
    Wheat = list(
        FrostLowTT = -4.0, FrostMaxReductionRatio = 0.4, FrostUpTT = 1.0, FrostMinReductionRatio = 0,
        FrostStartSensitiveGS = 6.4854544, FrostStartMostSensitiveGS = 8.0,
        FrostEndMostSensitiveGS = 9.4983302, FrostEndSensitiveGS = 9.4990961,
        HeatLowTT = 30.3459986, HeatMinReductionRatio = 0, HeatUpTT = 34.0000667, HeatMaxReductionRatio = 0.2031898,
        HeatStartSensitiveGS = 7.2832313, HeatStartMostSensitiveGS = 7.483232,
        HeatEndMostSensitiveGS = 8.9709024, HeatEndSensitiveGS = 9.0945434
    ),
    Canola = list(
        FrostLowTT = -4.108862, FrostMaxReductionRatio = 0.3050003, FrostUpTT = -0.7549879, FrostMinReductionRatio = 0,
        FrostStartSensitiveGS = 7.222864, FrostStartMostSensitiveGS = 7.329504,
        FrostEndMostSensitiveGS = 8.063452, FrostEndSensitiveGS = 8.181088,
        HeatLowTT = 27.45271, HeatMinReductionRatio = 0, HeatUpTT = 38.8354, HeatMaxReductionRatio = 0.5810167,
        HeatStartSensitiveGS = 5.578273, HeatStartMostSensitiveGS = 6.260291,
        HeatEndMostSensitiveGS = 9.200536, HeatEndSensitiveGS = 9.428058
    )
)

# create separate 2x2 figures (frost/heat × sensitivity/temperature) for each crop
for (crop in names(params)) {
    p <- params[[crop]]

    # local implementations (different names to avoid duplicating later helpers)
    frost_pot_fn <- function(t, low, up, maxR, minR) {
        out <- ifelse(t <= low, maxR,
                                    ifelse(t >= up, minR,
                                                 maxR + (t - low) * (minR - maxR) / (up - low)))
        pmin(pmax(out, 0), 1)
    }
    heat_pot_fn <- function(t, low, up, minR, maxR) {
        out <- ifelse(t <= low, minR,
                                    ifelse(t >= up, maxR,
                                                 minR + (t - low) * (maxR - minR) / (up - low)))
        pmin(pmax(out, 0), 1)
    }
    stage_sens_fn <- function(gs, s0, s1, e1, e0) {
        approx(x = c(s0, s1, e1, e0), y = c(0, 1, 1, 0), xout = gs, rule = 2)$y
    }

    temps_seq <- seq(-10, 40, by = 0.1)
    gs_seq    <- seq(0, 10, by = 0.01)

    pot_df_crop <- data.frame(
        temp = temps_seq,
        frost = frost_pot_fn(temps_seq, p$FrostLowTT, p$FrostUpTT, p$FrostMaxReductionRatio, p$FrostMinReductionRatio),
        heat  = heat_pot_fn(temps_seq, p$HeatLowTT, p$HeatUpTT, p$HeatMinReductionRatio, p$HeatMaxReductionRatio)
    )

    sens_df_crop <- data.frame(
        gs = gs_seq,
        frost = stage_sens_fn(gs_seq, p$FrostStartSensitiveGS, p$FrostStartMostSensitiveGS, p$FrostEndMostSensitiveGS, p$FrostEndSensitiveGS),
        heat  = stage_sens_fn(gs_seq, p$HeatStartSensitiveGS, p$HeatStartMostSensitiveGS, p$HeatEndMostSensitiveGS, p$HeatEndSensitiveGS)
    )

    # panels
    p_frost_sens <- ggplot(sens_df_crop, aes(gs, frost)) +
        geom_line(size = 1, color = "#1b9e77") +
        labs(x = "Growth stage (GS)", y = "Sensitivity (0–1)", title = "Frost: sensitivity") +
        theme_minimal() + ylim(0, 1) +
        xlim(c(p$FrostStartSensitiveGS - 0.5, p$FrostEndSensitiveGS + 0.5))

    p_heat_sens <- ggplot(sens_df_crop, aes(gs, heat)) +
        geom_line(size = 1, color = "#d95f02") +
        labs(x = "Growth stage (GS)", y = "Sensitivity (0–1)", title = "Heat: sensitivity") +
        theme_minimal() + ylim(0, 1) +
        xlim(c(p$HeatStartSensitiveGS - 0.5, p$HeatEndSensitiveGS + 0.5))

    p_frost_temp <- ggplot(pot_df_crop, aes(temp, frost)) +
        geom_line(size = 1, color = "#1b9e77") +
        labs(x = "Temperature (°C)", y = "Potential reduction (0–1)", title = "Frost: temperature response") +
        theme_minimal() + ylim(0, 1) +
        xlim(c(p$FrostLowTT - 2, p$FrostUpTT + 2))

    p_heat_temp <- ggplot(pot_df_crop, aes(temp, heat)) +
        geom_line(size = 1, color = "#d95f02") +
        labs(x = "Temperature (°C)", y = "Potential reduction (0–1)", title = "Heat: temperature response") +
        theme_minimal() + ylim(0, 1) +
        xlim(c(p$HeatLowTT - 2, p$HeatUpTT + 2))

    # assemble 2x2 (left = frost, right = heat; top = sensitivity, bottom = temperature)
    fig <- (p_frost_sens | p_heat_sens) / (p_frost_temp | p_heat_temp) +
        plot_layout(guides = "collect") & theme(legend.position = "none")
    print(fig + plot_annotation(title = crop))
}

```

## Methodology

All internal and output variables are reset at sowing and then updated daily during crop growth until harvest at event [DoManagementCalculations](/docs/Models/Events.html#ManagementCalculations).

### Daily reduction calculations

The daily frost ($R_{frost}$) and heat ($R_{heat}$) reductions are calculated by multiplying the potential reduction ratio (based on temperature) and the sensitivity (based on growth stage) for each stress.

$$
R_{frost} = P_{frost}(T_{min}) \times S_{frost}(GS)
$$

where:

- $P_{frost}(T_{min})$ is the potential reduction ratio from the frost potential reduction function above given the daily minimum temperature ($T_{min}$).
- $S_{frost}(GS)$ is the sensitivity from the frost sensitivity function above given the current crop growth stage (GS).

$$
R_{heat} = P_{heat}(T_{max}) \times S_{heat}(GS)
$$

where:

- $P_{heat}(T_{max})$ is the potential reduction ratio from the heat potential reduction function above given the daily maximum temperature ($T_{max}$).
- $S_{heat}(GS)$ is the sensitivity from the heat sensitivity function above given the current crop growth stage (GS).

The total daily reduction ratio from both stresses is then calculated as:

$$
R_{frost\_heat} = 1 - (1 - R_{frost}) \times (1 - R_{heat})
$$


The cummulative frost reduction from sowing to today (t) is calculated as:

$$
R_{frost,\ cum} (t) = 1 - \prod_{i=1}^{t} (1 - R_{frost} (i))
$$

The cumulative heat reduction from sowing to today (t) is calculated as:

$$
R_{heat,\ cum} (t) = 1 - \prod_{i=1}^{t} (1 - R_{heat} (i))
$$

The cumulative frost and heat reduction from sowing to today (t) is calculated as:
$$
R_{frost\_heat,\ cum} (t) = 1 - R_{frost,\ cum} (t)  \times R_{heat,\ cum} (t)
$$

Finally the frost- and heat-limited yield (Y_{frost\_heat}) at harvest is calculated as:

$$
Y_{frost\_heat} = Y \times (1 - R_{frost\_heat,\ cum})
$$

where:

- $Y$ is the potential yield (g/m²) estimated by the crop model without considering frost and heat damage.

## Events 


**Events Listened For**

These are signals or notifications that the function listens for from other parts of the simulation:

| Event          | Purpose                                                                                  |
|----------------|------------------------------------------------------------------------------------------|
| [Sowing](/docs/Models/Events.html#Sowing)   | Initialize internal variables for frost and heat damage functions |
| [DoManagementCalculations](/docs/Models/Events.html#ManagementCalculations)   | Perform the actual calculations for frost and heat damage functions |



**Events Raised to**

> No events are raised by this function.


## Properties


**Configurable and Reportable Properties**

> No configurable properties are available for this function.

**Read-Only Reportable Properties**

| Property | Type | Description                                                                                  |
|--------|---:|--------------------------------------------------------------------------------------------------|
| `FrostPotentialReductionRatio` | double | Daily potential yield reduction ratio by a frost event. |
| `FrostSensitivity` | double | Daily sensitivity of yield reduction to growth stage when the frost event occurs. |
| `FrostReductionRatio` | double | Daily actual yield reduction ratio by frost stress. |
| `HeatPotentialReductionRatio` | double | Daily potential yield reduction ratio by a heat event. |
| `HeatSensitivity` | double | Daily sensitivity of yield reduction to growth stage when the heat event occurs. |
| `HeatReductionRatio` | double | Daily actual yield reduction ratio by heat stress. |
| `FrostHeatReductionRatio` | double | Daily actual yield reduction ratio by combined frost and heat stress. |
| `CumulativeFrostReductionRatio` | double | Cumulative actual yield reduction ratio induced by frost stress. |
| `CumulativeHeatReductionRatio` | double | Cumulative actual yield reduction ratio induced by heat stress. |
| `FrostEventNumber` | double | Number of frost events during the sensitive period. |
| `HeatEventNumber` | double | Number of heat events during the sensitive period. |
| `CumulativeFrostHeatReductionRatio` | double | Cumulative actual yield reduction ratio induced by both frost and heat stress. |
| `FrostHeatYield` | double | Frost- and heat-limited yield (units: g/m²). |
| `FrostSensitivePeriodStartDAS` | double | Start of frost sensitive period in days after sowing (units: days). |
| `FrostSensitivePeriodEndDAS` | double | End of frost sensitive period in days after sowing (units: days). |
| `HeatSensitivePeriodStartDAS` | double | Start of heat sensitive period in days after sowing (units: days). |
| `HeatSensitivePeriodEndDAS` | double | End of heat sensitive period in days after sowing (units: days). |



## User Interface

`FrostHeatDamageFunctions` can be added as a child of a Crop Model in the model tree.

## Practical Example

Consider a wheat crop experiencing two frost events during the critical anthesis period:

- **Day 1 (Anthesis)**: Growth stage = 8.0, minimum temperature = -1°C
- **Day 2 (Post-anthesis)**: Growth stage = 8.2, minimum temperature = -2°C

Using the default wheat parameters from the table above, we can calculate the frost damage step by step.

#### Step 1: Calculate Potential Reduction Ratios

For each day, we calculate the potential reduction using the frost temperature response function:

**Day 1** (T_min = -1°C):
- Since -1°C is between FrostLowTT (-4°C) and FrostUpTT (1°C), we use linear interpolation:
- $P_{frost}(-1) = 0.40 + (-1 - (-4)) \times \frac{(0.00 - 0.40)}{(1 - (-4))} = 0.40 + 3 \times \frac{-0.40}{5} = 0.40 - 0.24 = 0.16$

**Day 2** (T_min = -2°C):
- $P_{frost}(-2) = 0.40 + (-2 - (-4)) \times \frac{(0.00 - 0.40)}{(1 - (-4))} = 0.40 + 2 \times \frac{-0.40}{5} = 0.40 - 0.16 = 0.24$

#### Step 2: Calculate Growth Stage Sensitivity

For both days, we calculate sensitivity using the frost sensitivity function:

**Day 1** (GS = 8.0):
- GS 8.0 is between FrostStartMostSensitiveGS (8.0) and FrostEndMostSensitiveGS (9.5)
- Therefore: $S_{frost}(8.0) = 1.0$ (maximum sensitivity)

**Day 2** (GS = 8.2):
- GS 8.2 is also between 8.0 and 9.5
- Therefore: $S_{frost}(8.2) = 1.0$ (maximum sensitivity)

#### Step 3: Calculate Daily Frost Reduction Ratios

**Day 1**: $R_{frost}(1) = P_{frost}(-1) \times S_{frost}(8.0) = 0.16 \times 1.0 = 0.16$ (16% yield loss)

**Day 2**: $R_{frost}(2) = P_{frost}(-2) \times S_{frost}(8.2) = 0.24 \times 1.0 = 0.24$ (24% yield loss)

#### Step 4: Calculate Cumulative Frost Damage

The cumulative frost reduction after both events is:

$$R_{frost,\ cum}(2) = 1 - \prod_{i=1}^{2} (1 - R_{frost}(i))$$

$$R_{frost,\ cum}(2) = 1 - [(1 - 0.16) \times (1 - 0.24)]$$

$$R_{frost,\ cum}(2) = 1 - [0.84 \times 0.76] = 1 - 0.6384 = 0.3616$$

#### Step 5: Calculate Final Yield Impact

If the potential wheat yield was 5000 kg/ha (500 g/m²), the frost-limited yield would be:

$$Y_{frost} = Y \times (1 - R_{frost,\ cum}) = 500 \times (1 - 0.3616) = 500 \times 0.6384 = 319.2 \text{ g/m²}$$

**Result**: The two frost events result in a cumulative yield loss of **36.2%**, reducing the potential yield from 5000 kg/ha to 3192 kg/ha.

This example demonstrates how:

1. Multiple frost events have multiplicative (not additive) effects
2. The timing during the sensitive period (anthesis) maximizes damage
3. Colder temperatures cause proportionally more damage
4. The cumulative effect (36.2%) is less than the sum of individual events (16% + 24% = 40%)

## See Also

* **Source code:** [FrostHeatDamageFunctions.cs on GitHub](http://github.com/APSIMInitiative/ApsimX/blob/master/Models/Functions/FrostHeatDamageFunctions.cs)
