---
title: "Haun Stage"
format: html
---

```{r setup, echo = FALSE}
setup_qmd()
```


Haun Stage is defined as the number of leaf tips that have appeared on a cereal plant since emergence. It estimates from daily thermal time and phyllochron.

## Overview

The Haun Stage model simulates the appearance of new leaf tips on cereal crops, such as wheat, from the time the plant emerges until the flag leaf appears [@haun_visual_1973]. It calculates from daily [thermal time](ThermalTime.qmd) and a cultivar-specific [phyllochron](Phyllochron.qmd), which is the thermal time required for each new leaf to appear.  


## Methodology

The Haun Stage is based on the biological principle that leaves appear at regular intervals, determined by temperature. Each day, the model calculates how much “thermal time” (accumulated heat) the plant has experienced. When enough thermal time has passed, a new leaf tip appears.

The daily increase in Haun Stage is calculated as:

$$
\Delta \text{HS} = \frac{\text{TT}}{\text{Ph}}
$$

Where:

- $\text{TT}$ is the amount of heat accumulated in a day (measured in degree-days).
- $\text{Ph}$ is the amount of thermal time needed for each new leaf to appear.

The model sums these daily increments from emergence until the flag leaf appears, for example, a Haun Stage of 5.3 means five leaves are fully visible and the sixth is partly emerged (30%).

**Key assumptions:**

- Leaf appearance is driven mainly by temperature.
- Each variety has its own characteristic phyllochron.
- The process is continuous, not stepwise, allowing for partial leaves.

This approach is supported by research showing that leaf appearance in cereals follows a predictable thermal pattern [@haun_visual_1973].

## Cultivar-Specific Parameters

> no cultivar-specific parameters are used in the Haun Stage model.

## Practical Example

Suppose a researcher wants to estimate when the flag leaf will appear in different wheat varieties. They set the phyllochron for each variety based on field data. For example, if a variety has a phyllochron of 120 °Cd and one day records 24 °Cd of thermal time:

$$
\Delta \text{HaunStage} = \frac{24}{120} = 0.2
$$

After five days at this rate, the Haun Stage increases by 1.0, meaning one new leaf tip has appeared. By tracking Haun Stage over time, researchers can compare development rates between varieties or under different treatments, supporting breeding and agronomic decisions.
The Haun Stage model tracks the number of leaf tips that have appeared on a cereal plant since emergence. It provides a continuous measure of crop development and is critical for predicting phenological stages and leaf-related processes.


## Simulation Example

We demonstrate progress of Haun Stage using [example simulations](/docs/Plants/Wheat/Examples/wheat_example1.qmd) for wheat. 

Two cultivars, `Emu_Rock` and `Sunlamb`, are used in the example simulations to represent the very quick and very late maturity types, respectively [@celestina_cultivar_2023]. Both cultivars have the similar [base phyllochron values](Phyllochron.qmd), resulting in similar Haun Stage.


```{r fig-huan-stage}
plot_output("docs/Plants/Wheat/Examples/wheat_example1.apsimx",
    "Wheat.Phenology.HaunStage", stage_name = TRUE,
    max_stage = 6)

```


