---
title: "Canola Example Simulation 1"
output: html_document
---


```{r setup, echo = FALSE}
setup_qmd()
```


This example demonstrates a [Canola](../index.qmd) simulation using APSIM Next Generation (APSIM NG) at two contrasting locations in Australia, two different cultivars, and two different sowing dates. The simulation is designed to illustrate the impact of these variables on canola growth and development.

## Simulation Configuration

The two contrasting cultivars are `GenericEearly` for early maturity type and `GenericLate` for late maturity type, which represent different growth characteristics. The two different sowing dates are `1st May` and `1st June`.

The two contrasting locations are Roma in Queensland and Hamilton in Victoria, Australia. The weather data are obtained from SILO Patched Point Dataset.


```{r sites}
sites <- file.path("_resources/mets", c("Roma.met", "Hamilton.met"))  |> 
    here::here() |> 
    weaana::readWeatherRecords() |> 
    weaana::siteInfor() |> 
    mutate(Site = c("Roma", "Hamilton"))

sites |> 
    ggplot() +
    myutils::geom_oz() +
    geom_point(aes(x = Longitude, y = Latitude)) +
    myutils::theme_map() +
    coord_fixed()

```
