---
title: "Heading"
format: html
---


```{r setup, echo = FALSE}
setup_qmd()
```

The **Heading** phase represents the developmental stage between `FlagLeafAppearance` and `HeadEmergence` in wheat crops. It models the progression of time and environmental responses that lead to the visible emergence of the crop's head.

## Overview

The [GenericPhase](/docs/Models/PMF/Phenology/Phases/GenericPhase.qmd) model simulates the time between the `FlagLeafAppearance` and `HeadEmergence` stages.

- The **Heading** simulates the transition from flag leaf appearance to head emergence.
- The model uses daily [Thermal Time](ThermalTime.qmd), retrieved from the `[Phenology].ThermalTime` function.
- The **Target thermal time** is depending on photoperiod response and fixed constant of thermal time target at long days
$$
\text{Target} = \text{LongPpBase} \times \left( 1 +  \text{PpSens} \times \text{PpResponse} \right)
$$

Where:

- $\text{LongPpBase}$ = base thermal time under long day condition, i.e. at 20 hours of daylight with default value of 200.0 °Cd
- $\text{PpSens}$ = sensitivity to photoperiod, with default value of 2 (unitless multiplier)
- $\text{PpResponse}$ = interpolated modifier (0 to 1) based on day length


The photoperiod response is modeled using a piecewise linear interpolation:

| Photoperiod (h) | Response |
|------------------|----------|
| 6.0              | 1.0      |
| 8.0              | 1.0      |
| 20.0             | 0.0      |
| 22.0             | 0.0      |

```{r photoperiod-effect, fig.height = 3, fig.cap="Photoperiod Effect on Heading Phase"}

# Plot the photoperiod effect function f_d(d) and f_pp(d)
library(ggplot2)
x <- c(6, 8, 20, 22)
y <- c(1, 1, 0, 0)
x_out <- seq(5, 22)
df <- data.frame(day_length = x_out, f_d = weaana::interpolationFunction(x, y, values = x_out))

ggplot(df, aes(x = day_length)) +
    geom_line(aes(y = f_d)) +
    labs(
        x = "Day Length (hours)",
        y = "Modifier Value"
    ) +
    scale_x_continuous(breaks = x_out) +
    theme_minimal()
```

The daily Photoperiod is calculated using the [PhotoperiodFunction](/docs/Models/Functions/PhotoperiodFunction.qmd) with 6 degrees twilight.

## Cultivar-Specific Parameters

These parameters can vary between crop cultivars and are typically found in the `Phenology` module:

| Name                                                  | Description                                                        | Default Value |
|-------------------------------------------------------|--------------------------------------------------------------------|---------------|
| [Phenology].HeadEmergenceLongDayBase.FixedValue       | Base thermal time (°Cd) required under long days                   | 200           |
| [Phenology].HeadEmergencePpSensitivity.FixedValue     | Sensitivity to photoperiod (unitless multiplier)                   | 2             |

## Practical Example

**Effect of Photoperiod Sensitivity on Heading Phase**
Let's recalculate the target thermal time for a day length of 12 hours. The photoperiod response $f_d(12)$ can be interpolated from the table above:

- Between 8 h (response = 1.0) and 20 h (response = 0.0), the response decreases linearly.
- $f_d(12) = 1.0 - \frac{12 - 8}{20 - 8} = 1.0 - \frac{4}{12} = 1.0 - 0.333 = 0.667$

Now, for each $\text{PpSens}$:

| Photoperiod Sensitivity ($\text{PpSens}$) | $1 + \text{PpSens} \times 0.667$ | Target Thermal Time (°C·days) |
|-------------------------------------------|----------------------------------|-------------------------------|
| 0.0                                       | $1 + 0.0 \times 0.667 = 1.000$   | $200 \times 1.000 = 200.0$    |
| 0.5                                       | $1 + 0.5 \times 0.667 = 1.333$   | $200 \times 1.333 = 266.7$    |
| 1.0                                       | $1 + 1.0 \times 0.667 = 1.667$   | $200 \times 1.667 = 333.3$    |
| 1.5                                       | $1 + 1.5 \times 0.667 = 2.000$   | $200 \times 2.000 = 400.0$    |
| 2.0                                       | $1 + 2.0 \times 0.667 = 2.333$   | $200 \times 2.333 = 466.7$    |


## See Also

* [Photoperiod Effect on Heading in Australian Wheat](/tips/heading-ppd-oz-wheat.qmd)
