---
title: "Fertiliser Types"
date: "2025-06-06"
output: html_document
---

```{r setup, echo = FALSE}

suppressPackageStartupMessages({
  library(rapsimng)
  library(tidyverse)
  library(knitr)
})
knitr::opts_chunk$set(warning = FALSE, 
                  echo = FALSE,
                  message = FALSE)
```
## Introduction

The fertiliser module defines the characteristics of different fertiliser types, such as their chemical composition and release behaviour. These definitions are stored in a resource file named [Fertiliser.json](https://raw.githubusercontent.com/APSIMInitiative/ApsimX/9c3f72cd2b7641de4f72463c427ccf3e5d85b8a3/Models/Resources/Fertiliser.json).

Understanding the composition and behaviour of different fertilisers allows users to select the most appropriate input for specific scenarios, such as rapid nitrogen availability, slow-release fertilisers, or dual-nutrient blends.

## Structure of Fertiliser Definitions

Each fertiliser type in APSIM has the following attributes:

- **Name**: Identifier used within the simulation to refer to this fertiliser.
- **Description**: A human-readable explanation of the fertiliser's nature or composition.
- **Solute**: The main chemical form of the nutrient (e.g., nitrate, ammonium) that the fertiliser releases into the soil.
- **Fraction**: The proportion (from 0 to 1) of the nutrient solute in the fertiliser product.
- **Release Rate**: The fixed amount or rate at which the nutrient becomes available in the soil after application. In most cases, this represents instantaneous release, but could be extended for controlled-release products.

## Example Fertiliser Types in APSIM

Below is a simplified summary of several fertiliser types defined in APSIM NG:


```{r read-data}


a <- read_apsimx("https://raw.githubusercontent.com/APSIMInitiative/ApsimX/9c3f72cd2b7641de4f72463c427ccf3e5d85b8a3/Models/Resources/Fertiliser.json")

fertilisers <- a$Children[[1]]$Children

fertilisers |> 
    map_df(function(x) {
        tibble(Name = x$Name,
               Description = x$Description,
               Solute1Name = x$Solute1Name,
               Solute1Fraction = x$Solute1Fraction,
               Release = x$Children[[1]]$FixedValue)
    }) |> 
  kable()


```
