---
title: "Thermal Time"
output: html_document
---


```{r setup, echo = FALSE}
setup_qmd()
```

Daily thermal time for wheat based on [SubDailyInterpolation](/docs/Models/Functions/SubDailyInterpolation.qmd) with three cardinal temperature and [ThreeHourAirTemperature](/docs/Models/Functions/ThreeHourAirTemperature.qmd) interpolation method.

## Overview

The **Thermal Time** function in wheat model calculates the daily thermal time. This is crucial for crop development and organ growth. The model uses a sub-daily interpolation method to estimate air temperatures at three hourly interpolation method ([ThreeHourAirTemperature](/docs/Models/Functions/ThreeHourAirTemperature.qmd)), then applies a three cardinal temperature function to each interpolated temperature. These values are averaged to give a daily  thermal time used in phenological and growth calculations.


* `InterpolationMethod`: [ThreeHourAirTemperature](/docs/Models/Functions/ThreeHourAirTemperature.qmd) is used for sub-daily interpolation, i.e. 3-hourly temperature estimates.
* `Response`: Three cardinal temperature using [XYPairs](/docs/Models/Functions/XYPairs.qmd) to define the temperature response function. This function determines how effective the temperature is for thermal time accumulation.

| Cardinal Point | Description                | Value (°C) |
|:--------------:|:--------------------------|:----------:|
| Base           | Minimum for growth         | 0          |
| Optimum        | Maximum rate of growth     | 26         |
| Maximum        | Growth ceases above this   | 37         |

```{r plot-thermal-time, fig.height = 3, fig.cap="Thermal Time Function with Three Cardinal Temperatures"}
library(ggplot2)


x <- c(0, 26, 37)
y <- c(0, 26, 0)
x_out <- seq(-5, 40)
df <- data.frame(Temperature = x_out, 
                    Response = weaana::interpolationFunction(x, y, values = x_out))


ggplot(df, aes(x = Temperature, y = Response)) +
    geom_line(color = "darkorange", size = 1.2) +
    geom_vline(xintercept = x, linetype = "dashed", color = "grey50") +
    annotate("text", x = x[1], y = max(df$Response), label = "Base", vjust = -0.5, hjust = -0.1, size = 3) +
    annotate("text", x = x[2] - 3, y = max(df$Response), label = "Optimum", vjust = -0.5, hjust = 0.5, size = 3) +
    annotate("text", x = x[3], y = max(df$Response), label = "Maximum", vjust = -0.5, hjust = 1.1, size = 3) +
    labs(
        x = "Temperature (°C)",
        y = "Thermal Response"
    ) +
    theme_minimal()

```

## Cultivar-Specific Parameters

> In general, we do not modify the thermal time function for different cultivars. 

## Practical Example

> No practical example is provided for the thermal time function.
