---
title: "Budding"
format: html
---

The **Budding** phase in [Chickpea](/docs/Plants/Chickpea) represents the developmental period from **StartBudding** to **StartFlowering**. This phase covers the period when floral buds are visible and developing until the first flowers open.

## Overview

The [GenericPhase](/docs/Models/PMF/Phenology/Phases/GenericPhase.qmd) model simulates the time between **StartBudding** and **StartFlowering** stages.

- This phase starts at **StartBudding** and ends at **StartFlowering**
- The model uses daily [Thermal Time](ThermalTime.qmd), retrieved from the `[Phenology].ThermalTime` function
- The **Target thermal time** is cultivar-specific and influenced by photoperiod at end of the juvenile phase.

## Bud Development

During the budding phase:

- Floral buds become visible at nodes
- Buds continue to develop and enlarge
- The plant continues some vegetative growth
- Branching may continue depending on environmental conditions and cultivar


The transition from bud development to flowering is primarily driven by thermal time accumulation and photoperiod at the end of the juvenile phase [@robertson_simulation_2002].

```{r tbl-target}
pd <- tibble::tibble(
    `Photoperiod` = c(10, 12.43, 17),
    `Target thermal time` = c(501.6, 50, 50)
)
pd |> 
    knitr::kable(col.names = c("Photoperiod (h)", "Target thermal time (°Cd)"), digits = c(1, 1),
                caption = "Photoperiod influence on target thermal time for budding phase in chickpea [@robertson_simulation_2002].")
```

```{r fig-target, fig.cap = "Photoperiod influence on target thermal time for budding phase in chickpea [@robertson_simulation_2002]."}
pd |> 
    ggplot2::ggplot(ggplot2::aes(x = Photoperiod, y = `Target thermal time`)) +
    ggplot2::geom_line() +
    ggplot2::geom_point() +
    ggplot2::labs(x = "Photoperiod (h)", y = "Target thermal time (°Cd)") +
    ggplot2::theme_minimal()
```

## Cultivar-Specific Parameters

| Name       | Description                                         | Default Value |
|------------|-----------------------------------------------------|---------------|
| [Phenology].Budding.Target.XYPairs.X   | Photoperiod at the end of the juvenile phase (h) | 10,12.43,17 |
| [Phenology].Budding.Target.XYPairs.Y   | Target thermal time for bud development (°Cd) | 501.6, 50, 50 |
